Final Activity Analysis
========================================================

Note: This analysis was created on a Mac. Some functions may be altered to be able to run on Windows machines.

-------
This markdown file was created to accompany the final alaysis write up and allow anyone to reproduce the results therein.

## Preliminaries

### Load the data and any required packages

```{r loadData}
load("../rawData/samsungData.rda")
library(randomForest)
library(caret)
```

### Explore the data and perform necessary transformations

```{r transformation, dependson = "loadData"}
# Check for missing values
sum(is.na(samsungData))
# Make sure column names are unique
length(unique(names(samsungData)))
# Create unique names
samsungData <- data.frame(samsungData)
# Change subject variable from numeric to factor
samsungData$subject <- as.factor(samsungData$subject)
# Change activity variable from character to factor
samsungData$activity <- as.factor(samsungData$activity)
```


No missing values, some columns were not unique. 

## Partition data into training, validation, and test sets.

```{r partition, dependson = c('loadData', 'transformation')}
samsung_training <- subset(samsungData, subject == 1 |
                             subject == 3 |
                             subject == 5 |
                             subject == 6 |
                             subject == 7 |
                             subject == 8)
samsung_training <- samsung_training[,c(1:561, 563)]
samsung_validation <- subset(samsungData, subject == 11 |
                               subject == 14 |
                               subject == 15 |
                               subject == 16)
samsung_validation <- samsung_validation[,c(1:561, 563)]
samsung_test <- subset(samsungData, subject == 27 |
                         subject == 28 |
                         subject == 29 | 
                         subject == 30)
```

## Make a model and figures

### Create a model using randomForest and predict on test set

```{r model, dependson = c('loadData', 'transformation', 'partition')}
samsung_forest <- randomForest(activity ~ ., data = samsung_training, prox = TRUE)
samsung_forest
predictions <- predict(samsung_forest, samsung_test)
# calculate failure rate
(sum(predictions != samsung_test$activity)/(length(predictions)))*100
```

### Create Figures
```{r figures, dependson= c('loadData','transformation','partition','model')}

# Figure showing the important variables
important <- samsung_forest$importance
important <- important[order(important[,1], decreasing = TRUE),]
important <- as.matrix(important)
plot(samsung_training[,rownames(important)[1]],
     samsung_training[,rownames(important)[3]],
     col = samsung_training$activity,
     xlab = rownames(important)[1],
     ylab = rownames(important)[3])

# Figure visualizing the errors
percents <- data.frame("actual" = c("laying", "sitting","standing","walk","waldown","walkup"),
                       "total mistakes" = c(0,48,34,13,12,14),
                       "laying" = c(0,0,0,0,0,0),
                       "sitting" = c(0,0,48,0,0,0),
                       "standing" = c(0,34,0,0,0,0),
                       "walking" = c(0,0,0,0,4,9),
                       "downstairs" = c(0,0,0,2,0,11),
                       "upstairs" = c(0,0,3,0,11,0))
barplot(as.matrix(percents[,3:8]), 
        col = percents$actual, 
        legend = percents$actual,
        xlab = "Actual Activity", 
        ylab = "Number of Errors",
        main = "Error Identification")
```
